---
title: "Get data for Paper Kartaram et al., "Citrulline"
author: "Marc A.T. Teunis"
date: "January 31, 2018"
output: html_document
---

## Aim
Selecting only those features that were used for the paper Kartaram et al: "Plasma citrulline concentration, a marker for intestinal functionality, reflects exercise intensity in healthy young men"

```{r, root, include=FALSE}
if(!require("rprojroot")) install.packages("rprojroot", dependencies = TRUE)
library(rprojroot)
root <- rprojroot::find_root_file(criterion = is_rstudio_project)
root
```

## Packages
```{r, packages}
library(tidyverse)
library(readr)
```

## read data
```{r, read_data}
data_file_path <- file.path(root, "data", "diagrams_clean.rds")
diagrams <- read_rds(path = data_file_path)
```

## Inspect data
```{r}
head(diagrams, 3)
# check factor levels
levels(diagrams$subject)
levels(diagrams$protocol)
levels(diagrams$time)
levels(diagrams$analyte %>% as.factor)
```

## Selecting analytes:
The analytes mentioned in the paper:

 - Glutamine (`gln`)
 - Alanine (`ala`)
 - Citrulline (`citrul`)
 - Arginine (`arg`)
 - Urea (`UREUM`)
 - IFABP (`ifabp`)
 - Cortisol (`CORT`)
 
The paper concerns only data from the 'GRINTA!' study 

```{r}
# create vector with analyte level-names
analytes_filter <- c("gln", "ala", "citrul", "arg", "UREUM", "ifabp", "CORT")
```
 
## Filter data 
The data is filtered for 'GRINTA!' and above mentioned analytes.

```{r, filter}
citrulline_data <- diagrams %>%
  filter(analyte %in% analytes_filter) %>%
  filter(study == "grinta") %>% droplevels(.)

# check factor levels
levels(citrulline_data$subject)
levels(citrulline_data$protocol)
levels(citrulline_data$time)
levels(citrulline_data$analyte %>% as.factor)
head(citrulline_data)
```

## Missing values
```{r}
sum(is.na(citrulline_data))
```

## Write data to disk
Data is written to disk as a tab separated value (*.txt) file named "./data/citrulline_paper_kartaram_et_al_1_0.txt". This data is used for creating the figures of the paper.
```{r}
write_tsv(citrulline_data, path = file.path(root, "data", "citrulline_paper_kartaram_et_al_1_0.txt"))
```

 
 